---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# gdpm

<!-- badges: start -->
<!-- badges: end -->

The goal of gdpm is to provide the data from the General Department of Preventive 
Medicine (GDPM) of Vietnam. 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("epix-project/gdpm")
```
## Example

```{r}
library(gdpm)
```

The package contains epidemiological data in 29 epidemiological data frames, each data frame corresponding to one syndromic disease. For example, `chickenpox`:

```{r}
chickenpox <-  getid(chickenpox)
head(chickenpox)
```

The data are expressed by the `incidence` or number of cases and `mortality` or the number of death per `province`, `month` and `year`.
See below, the section on the `getid` function, for an illustration of the specificity of this function.

The structure of the data frame is:

```{r}
str(chickenpox)
```

Note that time is coded by the 2 variables `year` and `month`. The latter is a factor in which the coding of the levels follows their chronological order:

```{r}
levels(chickenpox$month)
```

Head and tail of the data frame `chickenpox`:

```{r}
head(chickenpox)
tail(chickenpox)
```
