---
title: "Problem lecture"
csl: the-american-naturalist.csl
output:
  html_document:
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
<!-- bibliography: references.bib -->
---

```{r include=F}
knitr::knit_hooks$set(margin = function(before,options,envir) {
if(before) par(mgp=c(1.5,0.5,0),bty="n",plt=c(.105,.97,.13,.97)) else NULL })

knitr::opts_chunk$set(margin=T,prompt=T,comment="",collapse=T,cache=F,
dev.args=list(pointsize=11),fig.height=3.5,
fig.width=4.24725,fig.retina=2,fig.align="center")
```


```{r,include=FALSE}
library(magrittr) # for the " %>% " pipe
library(dplyr) # for "bind_rows", "rename", "mutate", "select"
cholera <- readRDS("cholera.RDS") %>% filter(year==1990)
hepatitis <- readRDS("hepatitis.RDS") %>% filter(year==1990)
error06 <- readRDS("error06.RDS")
error06 <- select(error06,-matches("disease")) %>% filter(duplicated(.)==FALSE)
```

## 1990 and the provinces


```{r}
unique(cholera$province)
unique(hepatitis$province)
grep(paste(unique(cholera$province),collapse = "|"),unique(hepatitis$province),value=T,invert=T)
grep(paste(unique(hepatitis$province),collapse = "|"),unique(cholera$province),value=T,invert=T)

```

Hypothesis : for **hepatitis and amoebiasis**, the number of province is 40 and we still have the "old-name" provinces. 

## 2006 and size of population

```{r}
pop_prlm <- filter(error06,duplicated(province)==TRUE)
pop_prlm$province
filter(error06,province==pop_prlm$province[1])
filter(error06,province==pop_prlm$province[2])
```



For **35 provinces**, we can see some differences in the size of the population between all the diseases and the sheet "RUBELLA". The size of the populations increases or decreases significantly for some provinces but for some of them, the difference is not big. 
We made some hypothesis to explain this difference: between 2006 and the other files, the order of the provinces has changed so maybe they forgot to also change the order of the population size. Also, some estimations of the size of the population are more precise in the other sheet disease than in rubella and sometimes it's the contrary.
We can also ask the question of **the data in RUBELLA**, are they also in the good or bad order like the provinces seem to be?


